---
title: "Q Statistics Error Propagation"
output: html_notebook
bibliography: bibliography.bib
---

We consider the formular for the Q-statistic as defined in @Keinan2009-ej.

\begin{equation}
    Q = \frac{\log(1 - 2E(F_\text{ST}^\text{auto}))}{\log(1 - 2E(F_\text{ST}^\text{X}))}
\end{equation}

Given errors $\Delta F_\text{ST}^\text{auto}$ and $\Delta F_\text{ST}^\text{X}$ we want to calculate $\Delta Q$, so we have

\begin{equation}
    \Delta Q = \sqrt{\left(
    \left(\Delta F_\text{ST}^\text{auto} \frac{\partial Q}{\partial F_\text{ST}^\text{auto}}\right)^2 +
    \left(\Delta F_\text{ST}^\text{X} \frac{\partial Q}{\partial F_\text{ST}^\text{X}}\right)^2
    \right)}
\end{equation}

with
\begin{equation}
    \frac{\partial Q}{\partial F_\text{ST}^\text{auto}} = 
    -\frac{2}{(1 - 2E(F_\text{ST}^\text{auto}))}
    \frac{1}{\log(1 - 2E(F_\text{ST}^\text{X}))}
\end{equation}

and
\begin{equation}
    \frac{\partial Q}{\partial F_\text{ST}^\text{X}} = 
    \frac{2}{(1 - 2E(F_\text{ST}^\text{X}))}
    \frac{\log(1 - 2E(F_\text{ST}^\text{auto}))}{(\log(1 - 2E(F_\text{ST}^\text{X})))^2}
\end{equation}

In our case we have $F_\text{ST}^\text{auto} = 0.003279 \pm 0.0001$ and $F_\text{ST}^\text{X} = 0.004018 \pm 0.00113$, which yields
```{r}
qstat <- function(fst_auto, fst_auto_err, fst_x, fst_x_err) {
  q <- log(1 - 2 * fst_auto) / log(1 - 2 * fst_x)
  deriv_auto <- - 2.0 / (1 - 2 * fst_auto) * 1 / log(1 - 2 * fst_x)
  deriv_x <- 2.0 / (1 - 2 * fst_x) * log(1 - 2 * fst_auto) / (log(1 - 2 * fst_x) ^ 2)
  qerr <- sqrt((fst_auto_err * deriv_auto) ^ 2 + (fst_x_err * deriv_x) ^ 2)
  return(c(q, qerr))
}
q1 <- qstat(0.003279, 0.0001, 0.004018, 0.00113)
q1
```

which yields a Z score of deviating from 0.75 of
```{r}
Z <- (q1[1] - 0.75) / q1[2]
Z
```
or a p-value of
```{r}
2*pnorm(-abs(Z))
```

For the numbers in @Keinan2009-ej, we have for example for the population pair "West African" vs. "North European", $F_\text{ST}^\text{auto} =  0.144 \pm 0.003$ and $F_\text{ST}^\text{X} = 0.221 \pm 0.009$, which yields a Q statistic of

```{R}
q2 <- qstat(0.144, 0.003, 0.221, 0.009)
q2
```

which gives a Z-score of deviating from 0.75 of
```{r}
Z <- (q2[1] - 0.75) / q2[2]
Z
```
and p-value of
```{r}
2 * pnorm(-abs(Z))
```

which isn't quite what's reported in @Keinan2009-ej (they report $p=3.0\times 10^{-8}$), but certainly still significant.



# References

